### 2011-10-06 doc/manuscript.tex
The manuscript is being rewritten.

Bacteria infect humans with a variety of diseases: e.g., food poisoning
(\textit{Eschericha Coli}), strep throat (\textit{Streptococcus pyrogenes}),
tuberculosis (\textit{Mycobacterium tuberculosis}), cholera (\textit{Vibrio
cholera}), anthrax (\textit{Bacillus anthracis}), pneumonia
(\textit{Chlamydophila pneumoniae}), and influenza (\textit{Haemophilus
influenzae}), to name a few.  \textit{Streptococcus pyogenes} (SPY) belonging to
group A streptococci (GAS) primarily infect humans in the throat and skin
\citep{Broyles2009}. It has high prevalence throughout the world causing
diseases ranging from mild illnesses such as pharyngitis to severe invasive
streptococcal diseases including necrotizing fasciitis and toxic shock syndrome
\citep{Cunningham2000a}.  Although \textit{Streptococcus dysgalactiae}
subspecies \textit{equisimilis} (SDE) belonging to group C or G streptococci
(GCS/GGS) used to be generally regarded as a veterinary pathogen
\citep{Vandamme1996}, it has been reported to cause serious and life-threatening
streptococcal diseases to humans traditionally associated with GAS
\citep[e.g.,][]{Brandt2009}.  Because of overlap in the disease profile and the
ecological niche of tissue sites betwen SPY and SDE, these bacteria may as well
interact with each other by exchanging parts of their genomes.

Therefore, the research topic on gene exchanges between SPY and SDE has been
discussed in various ways. \citet{Sachse2002} showed gene transfer from SDE to
SPY for the downstream of the genomic region of streptococcal pyrogenic
exotoxins. \citet{Davies2007} investigated phage 3396 from SDE as
a vehicle of gene transfer from SDE to SPY. \citet{Davies2009} demonstrated that
an integrative conjugative element of SDE allowed gene transfer from SDE to
other streptococci.  \citet{Ahmad2009} did not conclusively deduce the net
directionality of housekeeping gene transfer between SPY and SDE.  Notably,
\citet{Kalia2001} attempted to show that the direction of gene transfer from SPY
to SDE was more dominant than the other direction.  Because these researches on
the controversial issue, net directionality of gene transfer between SPY and
SDE, have been limited to relatively small regions in the genomes primarily on
representative loci from the Multi Locus Sequence Typing (MLST) database, it
remains to be elusive to determine the genome-wide degrees to which genomic
segments flowed between the two species, and net directions of thereof.

As sequencing costs plummet \citep{Mardis2011}, bacterial genome sequencing
produces individual genomes of high quality even within a bacterial species
\citep{Tettelin2009a}.  Studies on horizontal gene transfer in prokaryotes have
used genes in very large sets of species \citep{Koonin2001} from a great deal
of genomes available.  Although this type of study gave us insights into complex
evolutionary histories of prokaryotes, with multiple individual genomes per
species we now can go beyond the rough broad study on gene transfer towards
fine-scaled history of gene transfer between very closely related prokaryotes.
For instance, \citet{Liu2009} studied gene transfer events in
\textit{Lactobacillus bulgaricus} and \textit{Streptococcus thermophilus} using
the compositional approach developed by \citet{Karlin2001}. \citet{Luo2011}
investigated gene transfer events in \textit{Eschericha Coli} using the embedded
quartet decomposition analysis developed by \citet{Zhaxybayeva2006}.
\citet{Hamady2006} combined phylogenetic and compositional approaches to
detecting horizontal gene transfer.  \citet{Caro-Quintero2011} studied a number
of \textit{Shewanella baltica} genomes to identify the genetic elements that
enabled the species to adapt to redox gradients.  Although several complete
genomes from SPY have been available, \citet{Suzuki2011} recently added one more
complete genome of SDE to the previous set of a unique SDE genome
\citep{Shimomura2011}. The multiple SDE genomes prompted us to study
genome-wide gene flows between the two species that afflicted human beings
around the globe.

Bacteria evolves new biological features through horizontally transferred genes
from other lineages and vertically transmitted mutations on genes within the
same lineage.  Mutations on a gene modifiy relatively small portions of the gene
through substitutions, insertions, and deletions of deoxyribonucleic acids.
Horizontal gene transfers can reshape the organization of a bacterial genome,
which could revamp biological functions of the bacterial species.  Recipient
bacteria of transferred genes can substitute transferred genes for homologous
parts of their genomes, or integrate transferred genes into their genomes
without replacement. These two processes are here in this paper discriminated by
calling the former \textit{recombining gene transfer}, and the latter
\textit{horizontal gene transfer} \citep{Ochman2001,Lawrence2009}.  Despite the
two distinct evolutionary processes in bacteria, analyses of complete genome
sequences in efforts of better understanding bacterial evolution have paid less
attention to a combined approach to illuminating disparity of the two gene
transfer processes. We attempted here to use and compare two approaches for
distinguishing recombining gene transfer process from horizontal, or
non-recombining, gene transfer, and more systematically tackled their roles
in association with virulence genes.

Homologous recombining gene transfer could be studied more readily using the
core genomes from multiple individual bacterial genomes while horizontal gene
transfer might be more easily detected using the pan genomes. Recently, a method
of inferring homologous recombination with whole bacterial genomes was developed
\citep{Didelot2010}. The model of homologous recombination allowed them to study
flows of genetic segment between bacterial individuals of \textit{Bacillus
cereus} species in the core genomes.  More complex evolutionary processes acting
on bacterial genomes include genome rearragement, translocation, inversion,
duplication, gain, loss, etc.  Because of the complexity, these processes do not
readily lend themselves to statistically more sound model-based approaches such
as maximum likelihood. Instead, a parsimony-based approach to inferring gene
gain, loss, and transfer events was developed by  \citep{Doyon2011}.  We noted
that the first approach could be used to infer transfer of genomic segments due
to homologous recombination, and the second could do similarly attributed to
horizontal gene transfer.

In short, the complete genomes of the two spcies of SPY and SDE presented a
unique opportunity to study gene transfer over relatively recent time scales.
We used both of the model-based approach for inferring homologous recombination
and the parsimony-based one for detecting gene gain, loss, and horizontal
transfer. We also compared both of the results in recombining gene transfer
events to show that the parimony-based method inferred recombining gene transfer
events in similar biological functional categories to those inferred with the
model-based approach. This was the first attempt as far as we know to analyaze
SPY and SDE in genome-wide scale for detecting gene transfer events not only in
the core genome but also in the pan genome, which make the analysis complete in
terms of the amount of used genome data. Our analysis of core genomes also
showed for the first time that the net directionality of, at least, recombining
gene transfer from SPY to SDE was stronger than the reverse direction.

@@

\subsection{Recombination dirved more substitutions than did point mutation}

Five genomes except \textit{S.\ equi} ssp.\textit{\ equi} strain 4047 were
chosen (Table S\ref{tab:genome}) to study recombining gene transfer.  The genome
alignment of the five genomes showed two distinct genome structures of SPY and
SDE (Figure S\ref{fig:mauve}). The SDE genomes were larger in length than those
of SPY. Although it was unclear whether SDD was different from either of SDE1
and SDE2 from the graphical view of the genome alignment, the spcies tree
inferred from the genome alignment showed that SDD branch was attached to the
clade branch of SDE1 and SDE2 (Figure S\ref{fig:tree5}). The distribution of the
alignment block lengths was skewed to the right with mean of 4188 base pairs
long (Figure S\ref{fig:blocksize}), which implied that the model assumption of
geometric block length distribution might be plausible in \texttt{ClonalOrigin}
method (see section \textbf{Materials and Method} and \citealt{Didelot2010}).
The relative strength of recombination compared to mutations indicated that
reconmination-driven substitutions affected more sites than did mutation-driven
substitutions (Table S\ref{tab:clonalframe}). In the \textit{Streptococcus}
bacteria homologous recombination appeared to affect more sites than did point
mutations process.

@@

Along the genome of SDE1 population mutation rate did not appear to change much
(Figure S\ref{fig:scatter3}).  We found a few spots of genomic regions where the
mutation rates were larger than twice the global estimate of mutation rate.  The
recombination rate appeared to change more dynamically. Espeically, the genomic
region of SDE1 near 20 kilobase pairs from the origin of replication contained
alignment blocks with bell-shaped recombination rates.  The recombinant tract
length changes in a way that we could not note any particular positions along
the genome.  There appeared to be negative correlation between recombinant tract
length and recombination rate (Figure S\ref{fig:rhologdelta}).
% \texttt{WHAT ARE THESE BLOCKS? WHAT GENES ARE LOCATED IN THE BLOCK?}
% \texttt{USE UCSC GENOME BROWSER TO FIND WHAT GENES BELONG TO THOSE BLOCKS.}

@@

Most of local gene tree topologies (see section \textbf{Materials and Methods}),
or 66.7\%, along the alignment blocks were the same as the species tree, which
allowed us to be more confident of the species tree inferred using the alignment
blocks (Table S\ref{tab:Gene-tree-topologies}).  The second most probable
topology (9.3\%) was one where SDD was the outgroup, and SPY and SDE were
sister clades. SPY and SDE appeared to share about tenth of the core genomes that
they were more closely related with each other than each to SDD's genome. The
other topologies had less than 4\% of proportions.

\subsection{Unbalaned gene flow from SPY to SDE in genome-wide scale}

The heat map of Figure \ref{fig:Heatmap-of-recombination} (see Table
S\ref{tab:heatmap} for the values) shows the log base 2 ratios of the \textit{a
posteriori} number of recombinant edges for a pair of donor and recipient
species tree branches and their corresponding expected values given the species
tree and the three population parameters.  Because recombinant edges from any
species tree branch to the root branch are not possible under
\texttt{ClonalOrigin} model (see Figure S\ref{fig:clonalorigin}), the column for
the root as recipient was colored green, indicating invalid recombinant edges.
Similarly, all of the cells where a recipient branch was older than the donor,
and two branches did not share any contemporary time were colored green. The
diagonal cells were colored gray, indicating no significant number of
recombinant edges, because the observed and expected numbers of recombinant
edges were similar. As recombinant edges starting from and ending in the same
species tree branch would not change the species tree, which would not change
the likelihood much, the cells would be colored gray.  Because SDD is a strictly
veterinary pathogenic bacteria, we did not expect that SDD received much gene
transfer from either SPY1 or SPY2: the log base 2 ratio from SPY1 to SDD was
-2.20, and that from SPY2 to SDD was -1.97.  On the contrary, the recombing gene
transfer from SDE1 to SDD was 0.756, and that from SDE2 to SDD was 0.698. This
relatively higher recombining gene transfer from two SDE individuals to SDD
compared with that from two SPY individuals to SDD reflected that the SDD was
more closely related to SDE than to SPY in most of the genomes.

The genomic organization between SDD and SDE1/SDE2 were much similar (Figure
S\ref{fig:mauve}), and they formed a clade in the species tree
(S\ref{fig:tree5}). However, the gene transfer from SDD to either SDE1 or SDE2
was less frequent than the case of reverse direction; SDD served as a recipient
rather than a donor. It is not unclear why SDD received more genes from SDE1 or
SDE2 than did SDE1 or SDE2 from SDD.

Because it was expected that the recombining gene transfer within a clade would
be larger than that between clades, the recombining gene transfer log base 2
ratios between SDE1 and SDE2 (1.17 and 1.20) were about twice as large as that
from SDE1 and SDE2 to SPY1 and SPY2 (0.67, 0.66, 0.60, and 0.39).  The
intensities of recombining gene transfer from external branches of SDE to
external branches to SPY was similar to those from SDE1 or SDE2 to SDD (0.76 and
0.70).  Either SPY1 or SPY2, and SDD appeared to be no different from the view
point of SDE1 or SDE2 since the similar values of recombining gene transfer
intensity from SDE1 or SDE2 to the other three external branches.  Surprisingly,
the recombining gene transfer log base 2 ratios between SPY1 and SPY2 (1.06 and
1.07) was comparable with those from SPY1 and SPY2 to SDE1 and SDE2 (1.25, 1.25,
1.17, and 1.29).  Effectively, the recombining gene transfer from a clade SPY1
or SPY2 to a different clade SDE1 or SDE2 seemed to be no different from that
within the clade of SPY1 and SPY2. The two external branches of SDE received
genes from either SPY1 or SPY2 via recombining gene transfer as if SDE1 and SDE2
were in the same clade as SPY1 and SPY2.  The bacterial recombination between
SPY and SDE might happen more often than that between SDE and SDD even though
the genome configurations of SDE and SDD are more similar than each to SPY. This
might be consistent with the fact that SPY and SDE shared the same ecological
niche.  The gene transfer log base 2 ratio from SPY1 and SPY2 to SDE branch was
larger than any type of gene transfer: 1.88 and 1.94, respectively.  
Furthermore, we found a total of 58 genes for which a recombinant edge of
certain types were sampled with probability at least 0.9 by focusing on specific
types of recombinant edges.  (Table S\ref{tab:genes-transfer}).  Many of the
genes in the list were transferred from SPY to SDE rather than the reverse
direction overall.  All things considered, this clearly showed that the
recombining gene transfer was biased toward the net direction from SPY to SDE.

Figure \ref{fig:hgt-heatmap} shows the number of inferred non-recombining
horizontal gene transfer events between the different branches of the species
tree.  SDE1 and SDE2 shared the most events (26 and 34) compared to other sample
pairs.  SDD appeared to acquire many genes from the ancestral SDE branch.
Transfer events between the SPY and SDE clades were inferred in smaller numbers
(<6) compared to those within each clade (23--34). Among the clade-crossing
transfers, we did not see any significant enrichment for a specific direction,
as opposed to the result for recombining transfers from the model-based analysis
(Figure S\ref{fig:mowgli-recomb-heatmap}).

@@

Among the 2314 gene families the largest family consisted of 59 genes, while
nearly half of gene families (1066) had exactly one gene per species, or six
genes total (Figure S\ref{fig:famsize}).  Figure \ref{fig:Gene-duplication-loss}
shows the inferred gene counts for each extant and ancestral species. Overall,
we inferred 304 duplications, 944 recombining gene transfers, and 294
non-recombining gene transfers, comprising 23.7\% of total transfer events.  We
found that duplications made up only a small fraction of the events within the
clade (10.2\%), while horizontal transfers and losses were significantly more
frequent (41.7\% and 48.0\%, respectively). We found that duplications were most
often species-specific. Many gene losses likely represented horizontal transfers
that replaced the native gene sequence.  The species SDD and SDE were the most
expanded, both in terms of duplications and in-bound transfers.

@@

\texttt{Find cases of any virulence genes that were transferred via horizontally.} 
\texttt{Make a list of virulence genes and their categories with more interpretations.}

About 10\% of the genes comprised of virulence facters, about half of which were
found in the core genome of the five genome alignment. Using the UCSC genome
browser tracks we located virulence genes with notable posterior recombination
probability.  Stressful host conditions include heat shock, starvation, and iron
depletion.  Iron uptake of bacteria was known to be essential for survival of
the bacteria under hostile host condition. We found that parts of the gene,
putative calcium transporter (MgtA), in SDE1 had been transferred from SPY1
(SpyM3\_0440).  An extracellular protein, putative dipeptidase (SpyM3\_0465),
was found to be transferred from SPY to SDE. The gene as a whole was
transferred, which was a case for the finding that virulence genes tend to be
transferred as a whole unit \citep{Chan2009}. Dipeptidase may be important for
the survival of bacteria to scavenge dipeptides from the host.  An adhesins
gene, putative internalin A precursor (SpyM3\_1035), was found to be affected by
recombining gene transfer from SPY to SDE. Some of the middle parts of the gene
was affected.  A middle part of a chaperonin GroEL, putative heat shock protein
(SpyM3\_1765), was inferred to transfer from SPY to SDE1.  Increase of GroEL
expression was known to allow for the bacteria to survive the hostile host
environment of acid shock.  Superoxide dismutase was reported to be associated
virulence of \textit{Streptococcus agalactiae} \citep{Poyart2001}. This whole
gene including its 5' UTR, (SpyM3\_1071), was affected by recombining gene
transfer from SPY to SDE.  Pyruvate formate lyase was recently reported to have
virulence \citep{Yesilkaya2009}. A few parts of the gene (SpyM3\_1749) seemed to
be affected by recombining gene transfer.  One MLST locus, called DNA mismatch
repair protein MutS, appeared to be affected by recombining gene transfer.  A
putative spermidine/putrescine ABC transporter ATP-binding protein,
(SpyM3\_0764), was found to be transferred from SPY to SDE.  Under starvation
condition bacteria must use more effective energy metabolism.  A putative
glucose kinase, (SpyM3\_1180), might be transferred from SPY to SDE to adapt
SDE's energy metabolism to survive such harsh condition.

\subsection{Horizontal gene transfer is associated with virulence genes}

\texttt{Select genes with high horizontal gene transfers. Relate these to
categories of virulence genes. I should find many more virulence genes in Matt's
analysis. Use the virulence gene types from SDE2's paper. Which virulence gene
types have more gene transfer?}

We found that some virulence genes tended to have more horizontally transfers
than others. 
Streptococcal C5a peptidase (\textit{scpB}) (SDEG\_0933) was inferred to have six
horizontal gene transfer. 
Putative transposase were also found to have relatively large number of
horizontal gene transfer events.
(SpyM3\_1016, SpyM3\_1053, SpyM3\_1054)

Putative response regulator (SpyM3\_0174).
Putative maltose operon transcriptional repressor (SpyM3\_0982).
Putative histidine kinase (SpyM3\_0171).
Putative cell wall hydrolase (SpyM3\_0731, SpyM3\_0922, SpyM3\_1208).

GGS\_124 possesses 5 genes that encode putative nucleases with a secretion
signal peptide: genome-encoded streptodornase (SDEG\_0541), others,
prophage-associated deoxyribonuclease (SDEG\_1103), all of which are predicted to code for a secretion signal peptide.
Streptodornase (Sdn) - phage associated  (SpyM3\_1409).

Protein F2 like fibronectin-binding protein (SpyM3\_0104).

>>>>

Bacteria infect humans with a variety of diseases: e.g., food poisoning
(\textit{Eschericha Coli}), strep throat (\textit{Streptococcus pyogenes}),
tuberculosis (\textit{Mycobacterium tuberculosis}), cholera (\textit{Vibrio
cholera}), anthrax (\textit{Bacillus anthracis}), pneumonia
(\textit{Chlamydophila pneumoniae}), and influenza (\textit{Haemophilus
influenzae}), to name a few.  \textit{Streptococcus pyogenes} (SPY) belonging to
group A streptococci (GAS) primarily infect humans in the throat and skin
\citep{Broyles2009}. It has high prevalence throughout the world causing
diseases ranging from mild illnesses such as pharyngitis to severe invasive
streptococcal diseases including necrotizing fasciitis and toxic shock syndrome
\citep{Cunningham2000a}.  Its close relative, \textit{Streptococcus
dysgalactiae} subspecies \textit{equisimilis} (SDE) belonging to group C or G
streptococci (GCS/GGS), used to be generally regarded as a veterinary pathogen
\citep{Vandamme1996}. However, it has been reported to cause serious and
life-threatening streptococcal diseases to humans traditionally associated with
GAS \citep[e.g.,][]{Brandt2009}.  Because of the overlap in disease profile and
ecological niche of tissue sites betwen SPY and SDE, these bacteria may as
well interact with each other by exchanging parts of their genomes.

Therefore, the research topic on gene exchange between SPY and SDE has been
performed in various ways. \citet{Sachse2002} showed gene transfer from SDE to
SPY for the downstream of the genomic region of streptococcal pyogenic
exotoxins. \citet{Davies2007} investigated phage 3396 from SDE as a vehicle of
gene transfer from SDE to SPY. \citet{Davies2009} demonstrated that an
integrative conjugative element of SDE allowed gene transfer from SDE to other
streptococci.  \citet{Ahmad2009} did not conclusively deduce the net
directionality of housekeeping gene transfer between SPY and SDE.  Notably,
\citet{Kalia2001} attempted to show that the net direction of gene transfer from
SPY to SDE was more dominant than the other direction.  Because these researches
on the controversial issue, net directionality of gene transfer between SPY and
SDE, have been limited to relatively small regions in the genomes primarily on
representative loci from the Multi Locus Sequence Typing (MLST) database, it
remains to be elusive to determine the genome-wide degrees to which genomic
segments flowed between the two species, and net directions of thereof.

As sequencing costs plummet \citep{Mardis2011}, bacterial genome sequencing
produces individual genomes of high quality even within a bacterial species
\citep{Tettelin2009a}.  Studies on horizontal gene transfer in prokaryotes have
used genes in very large sets of species \citep{Koonin2001} from a great deal of
genomes available.  Although this type of study gave us insights into deep
evolutionary histories of prokaryotes, with multiple individual genomes per
species we now can go beyond the rough broad study on gene transfer towards
fine-scaled history of gene transfer between very closely related prokaryotes.
For instance, \citet{Liu2009} studied gene transfer events in
\textit{Lactobacillus bulgaricus} and \textit{Streptococcus thermophilus} using
the compositional approach of \citet{Karlin2001}. \citet{Luo2011} investigated
gene transfer events in \textit{Eschericha Coli} using the embedded quartet
decomposition due to \citet{Zhaxybayeva2006}.  \citet{Hamady2006} combined
phylogenetic and compositional approaches to detecting horizontal gene transfer.
\citet{Caro-Quintero2011} studied a number of \textit{Shewanella baltica}
genomes to identify the genetic elements that enabled the species to adapt to
redox gradients.  Although several complete genomes from SPY have been
available, \citet{Suzuki2011} just recently added one more complete genome of
SDE to the set of a previously unique SDE genome \citep{Shimomura2011}. These
multiple SDE genomes prompted us to study genome-wide gene flows between SPY and
SDE that afflicted human beings around the globe.

Bacteria evolves new biological features through horizontally transferred genes
from other lineages and vertically transmitted mutations on genes within the
same lineage.  Mutations on a gene modifiy relatively small portions of the gene
through substitutions, insertions, and deletions of deoxyribonucleic acids.
Gene transfers can reshape the organization of a bacterial genome,
which could revamp biological functions of the species.  Recipient bacteria of
transferred genes can substitute transferred genes for homologous parts of their
genomes, or integrate transferred genes into their genomes without replacement.
These two processes are here in this manuscript discriminated by calling the
former \textit{recombining gene transfer}, and the latter \textit{horizontal
gene transfer} \citep{Ochman2001,Lawrence2009}.  Despite the two distinct
evolutionary processes in bacteria, analyses of complete genome sequences in
efforts of better understanding bacterial evolution have paid less attention to
a combined approach to illuminating disparity of the two gene transfer
processes. We attempted here to use and compare two approaches for
distinguishing recombining gene transfer process from horizontal, or
non-recombining, gene transfer, and tackled their roles in association with
virulence genes.

Homologous recombining gene transfer could be studied more readily using the
core genomes of individual bacteria while horizontal gene transfer might be more
easily detected using the pan genomes. Recently, a method of inferring
homologous recombination with whole bacterial genomes was developed
\citep{Didelot2010}. The model of homologous recombination allowed them to study
flows of genetic segment between bacteria individuals of \textit{Bacillus
cereus} species in the core genomes.  Other evolutionary processes acting on
bacterial genomes include genome rearragement, translocation, inversion,
duplication, gain, loss, etc.  These processes did not readily lend themselves
to statistically more sound model-based approaches such as maximum likelihood.
Instead, a parsimony-based approach to inferring gene gain, loss, and transfer
events was developed by \citep{Doyon2011}.  We noted that the first approach
could be used to infer transfer of genomic segments due to homologous
recombination, and the second could do similarly attributed to horizontal gene
transfer.

In short, the complete genomes of the two spcies of SPY and SDE presented us a
unique opportunity to study gene transfer over relatively recent time scales.
We took advantage of the model-based approach for inferring homologous
recombination, and the parsimony-based one for detecting gene duplication, loss,
and horizontal transfer. We compared both of the results in recombining gene
transfer events to show that the parimony-based method inferred recombining gene
transfer events in biological functional categories similar to those inferred
with the model-based approach. This was the first attempt as far as we know to
analyaze SPY and SDE in genome-wide scale for detecting gene transfer events not
only in the core genome but also in the pan genome, which makes the analysis
complete in terms of the amount of used genome data. Our analysis of core
genomes also showed for the first time that the net directionality of
recombining gene transfer from SPY to SDE was stronger than the reverse
direction.

@@

\subsection{Recombination drived more substitutions than did point mutation}

Five genomes except \textit{S.\ equi} ssp.\textit{\ equi} strain 4047 were
chosen (Table S\ref{tab:genome}) to study recombining gene transfer using the
model-based approach.  The genome alignment of the five genomes showed two
distinct genome structures of SPY and SDE (Figure S\ref{fig:mauve}). The SDE and
SDD genomes were larger in length than those of SPY. Although it was unclear
whether SDD was different from either of SDE1 and SDE2 from the graphical view
of the genome alignment, the spcies tree inferred from the genome alignment
showed that SDD branch was attached to the clade branch of SDE1 and SDE2 (Figure
S\ref{fig:tree5}). The distribution of the alignment block lengths was skewed to
the right with mean of 4188 base pairs long (Figure S\ref{fig:blocksize}), which
supported that the model assumption of geometric block length distribution might
be plausible in \texttt{ClonalOrigin} method (see section \textbf{Materials and
Methods} and \citealt{Didelot2010}).  The relative strength ($r/m$) of
recombination compared to mutations indicated that recombination-driven
substitutions affected more sites than did mutation-driven substitutions (Table
S\ref{tab:clonalframe}). This could imply that in the \textit{Streptococcus}
species bacterial recombination process appeared to accelerate molecular
evolution of more sites than did point mutations process.

@@

Along the genome of SDE1 population mutation rate did not appear to change much
(Figure S\ref{fig:scatter3}).  We found a few spots of genomic regions where the
mutation rates were larger than twice the global estimate of mutation rate.  The
recombinant tract length changed in a way that we could not note any particular
positions with notable tract length values compared with other positions along
the genome.  As recombinant tract lengths were smaller, recombination rates per
site seemed larger (Figure S\ref{fig:rhologdelta}).  The recombination rate
appeared to change more dynamically. Especially, a genomic region of SDE1 near
20 kilobase pairs from the origin of replication contained alignment blocks with
bell-shaped recombination rates.

@@

There are 105 possible rooted topologies relating five taxa.  We expected the
dominant topology across sampled genealogies to be concordant with that of the
species tree.  Indeed, 66.7\% of the local gene tree topologies (see section
\textbf{Materials and Methods}) along the alignment blocks were the same as the
species tree, which allowed us to be more confident of the species tree inferred
using the alignment blocks (Table S\ref{tab:Gene-tree-topologies}).  Other
frequently sampled topologies were expected to be potentially informative about
common recombination events.  The second most frequent topology (9.3\%) was one
where SDD was the outgroup, and SPY and SDE were sister clades. Therefore, SPY
and SDE appeared be more closely related with each other than each to SDD's
genome in about tenth of the core genome. The other topologies had less than 4\%
of proportions.

\subsection{Unbalaned recombining gene flow from SPY to SDE in genome-wide scale}

The heat map of Figure \ref{fig:Heatmap-of-recombination} (see Tables
S\ref{tab:heatmap} and S\ref{tab:obsheatmap} for the heat map values and the
observed number of recombinant edges) shows the log base 2 ratios of the
\textit{a posteriori} number of recombinant edges for a pair of donor and
recipient species tree branches and their corresponding expected values given
the species tree (Figure S\ref{fig:tree5}) and the three population parameters
(Table S\ref{tab:three}).  Because recombinant edges from any species tree
branch to the root branch are impossible (see Figure S\ref{fig:clonalorigin}),
the column for the root as recipient was colored green, indicating invalid
recombinant edges.  Similarly, all of the cells for which a recipient branch was
older than the donor, and two branches did not share any contemporary time were
colored green. The diagonal cells were colored gray, indicating no significant
number of recombinant edges, because the observed and expected numbers of
recombinant edges were similar. As recombinant edges starting from and ending in
the same species tree branch would not change the species tree, which would not
change the likelihood much, the cells would be colored gray.  Because SDD is a
strictly veterinary pathogenic bacteria, we did not expect that SDD received
much gene transfer from either SPY1 or SPY2: the log base 2 ratio from SPY1 to
SDD was -2.20, and that from SPY2 to SDD was -1.97.  On the contrary, the
recombing gene transfer from SDE1 to SDD was 0.756, and that from SDE2 to SDD
was 0.698. This relatively higher recombining gene transfer from two SDE
individuals to SDD compared with that from two SPY individuals to SDD reflected
that the SDD was more closely related to SDE than to SPY in most of the genomes,
and the genomic organization between SDD and SDE1/SDE2 were much similar
(Figure S\ref{fig:mauve}).  However, the gene transfer from SDD to either SDE1
or SDE2 was less frequent than the case of reverse direction; SDD effectively
served as a recipient rather than a donor. It was not unclear why SDD received
more genes from SDE1 or SDE2 than did SDE1 or SDE2 from SDD.

Because it was expected that the recombining gene transfer within a clade would
be more frequent than that between clades, not surprisingly the recombining gene
transfer log base 2 ratios between SDE1 and SDE2 (1.17 and 1.20) were about
twice as large as that from SDE1 and SDE2 to SPY1 and SPY2 (0.67, 0.66, 0.60,
and 0.39).  The intensities of recombining gene transfer from external branches
of SDE1 and SDE2 to external branches of SPY1 and SPY2 was similar to those from
SDE1 or SDE2 to SDD (0.76 and 0.70).  After all, either SPY1 or SPY2, and SDD as
recipient appeared to be no different from the view point of SDE1 or SDE2 as
donor since the similar values of recombining gene transfer intensity from SDE1
or SDE2 to the other three external branches.  Surprisingly, the recombining
gene transfer log base 2 ratios between SPY1 and SPY2 (1.06 and 1.07) was
comparable with, but even less than those from SPY1 and SPY2 to SDE1 and SDE2
(1.25, 1.25, 1.17, and 1.29).  Effectively, the recombining gene transfer from a
clade SPY1 or SPY2 to a different clade SDE1 or SDE2 seemed to be no different
from that within the clade of SPY1 and SPY2.  The two external branches of SDE
received genes from either SPY1 or SPY2 via recombining gene transfer as if SDE1
and SDE2 were in the same clade as SPY1 and SPY2.  This was more surprising
because the bacterial recombination between SPY and SDE seemed to happen more
often than that between SDE and SDD even though the genome configurations of SDE
and SDD were more similar than each to SPY.  However, this result could be
concordant with the fact that SPY and SDE shared the same ecological niche
despite their different genomic organization.  The gene transfer log base 2
ratio from SPY1 and SPY2 to SDE branch was also larger than any type of
recombining gene transfer: 1.88 and 1.94, respectively.  Additionally we found
58 genes for which a recombinant edge of certain types were sampled with
probability at least 0.9 by focusing on specific types of recombinant edges
(Table S\ref{tab:genes-transfer}).  Many of the genes in the list were
transferred from SPY to SDE rather than the reverse direction overall.  All
things considered, this epitomized that the recombining gene transfer was biased
toward the net direction from SPY to SDE.

We tried in vain to quantify differential horizontal gene transfer between SPY
and SDE with the parsimony-based approach (Figure \ref{fig:hgt-heatmap}).  Among
the clade-crossing transfers, we did not see any significant enrichment for a
specific direction, as opposed to the result for recombining transfers from the
model-based analysis (Figure S\ref{fig:mowgli-recomb-heatmap}).  SDE1 and SDE2
shared the most events (26 and 34) compared to other sample pairs.  SDD appeared
to acquire many genes from the ancestral SDE branch.  Transfer events between
the SPY and SDE clades were inferred in smaller numbers (<6) compared to those
within each clade (23--34). We were puzzled by this finding because virulence
genes were found to be associated with horizontal gene transfer, and we could
not show net directionality of horizontal gene transfer between SPY and SDE.

@@

Overall, we inferred 304 duplications, 944 recombining gene transfers, and 294
horizontal gene transfers, comprising 23.7\% of total transfer events (Figure
\ref{fig:Gene-duplication-loss}).  Among the 2314 gene families the largest
family consisted of 59 genes, while nearly half of gene families (1066) had
exactly one gene per species, or six genes total (Figure S\ref{fig:famsizes}).
Within clades we found that duplications made up only a small fraction of the
events  (10.2\%), while horizontal transfers and losses were significantly more
frequent (41.7\% and 48.0\%, respectively). Duplications were most often
species-specific.  The species SDD and SDE were the most expanded, both in terms
of duplications and in-bound transfers.

@@

About 10\% of the genes comprised of virulence facters, about half of which were
found in the core genome of the five genome alignment. Using the UCSC genome
browser tracks we located virulence genes with notable posterior recombination
probability.  These genes were by no means comprehensive list of virulence
factors that might have been transferred via homologous recombination.
Stressful host conditions for pathogeneic bacteria can include heat shock,
starvation, and iron depletion.  Iron uptake of bacteria was known to be
essential for survival of the bacteria under hostile host condition. We found
that parts of putative calcium transporter (MgtA) in SDE1 could have been
transferred from SPY1 (SpyM3\_0440).  An extracellular protein, putative
dipeptidase (SpyM3\_0465) that could be important for the survival of bacteria
to scavenge dipeptides from the host, was inferred to be transferred from SPY to
SDE: the whole gene was transferred, which was a case for the finding that
virulence genes tended to be transferred as a whole unit \citep{Chan2009}.
Another example of whole gene transfer was superoxide dismutase that was
reported to be associated virulence of \textit{Streptococcus agalactiae}
\citep{Poyart2001}: the whole gene including its 5' UTR, (SpyM3\_1071), was
affected by recombining gene transfer from SPY to SDE.  Some of the genes were
only affected in their parts.  The middle parts of an adhesins gene, putative
internalin A precursor (SpyM3\_1035), was found to be affected by recombining
gene transfer from SPY to SDE.  The middle part of a chaperonin GroEL, putative
heat shock protein (SpyM3\_1765), was inferred to transfer from SPY to SDE1:
increase of GroEL expression was known to allow for the bacteria to survive the
hostile host environment of acid shock.  Parts of the gene (SpyM3\_1749),
pyruvate formate lyase that was recently reported to have virulence
\citep{Yesilkaya2009}, seemed to be affected by recombining gene transfer.
Under starvation condition bacteria must use more effective energy metabolism: a
putative glucose kinase, (SpyM3\_1180), were inferred to be transferred from SPY
to SDE to adapt SDE's energy metabolism to survive such harsh condition.  We
also found that an MLST locus, called DNA mismatch repair protein (MutS),
appeared to be affected by recombining gene transfer.

\subsection{Virulence genes via horizontal gene transfer}

Horizontal gene transfers were found to be significantly associated with
virulence genes (P-value of $<1.33 \times 10^{-11}$) whereas recombining gene
transfers were not enriched for virulence genes (P-value of $>0.90$) using the
parsimony-based approach.  This was also consistent with the fact that we could
not find the association of recombining gene transfer with virulence genes in
the model-based approach.  This supported more evidence that non-recombining
gene transfer served as a more effective means by which SDE or SPY acquired
virulence genes, compared to recombining gene transfer.  
Families with high numbers of duplications were also significantly enriched for
virulence genes (P-value of $<7.13 \times 10^{-3}$) whereas families with high
losses showed no enrichment (P-value of $>0.54$).

Roughly, half of the virulence genes were in the pan genome.  We did not attempt
to comprehensively list virulence genes that were inferred to be transferred
horizontally. We rather sort virulence genes by the number of horizontal gene
transfer events to list the first several genes with high ranks.  Streptococcal
C5a peptidase (\textit{scpB}, SDEG\_0933) was inferred to have six horizontal
gene transfer.  Putative transposase were found to have a relatively large
number of horizontal gene transfer events (SpyM3\_1016, SpyM3\_1053,
SpyM3\_1054). Other genes with relatively large number of horizontal gene
transfer events included putative response regulator (SpyM3\_0174), putative
maltose operon transcriptional repressor (SpyM3\_0982), putative histidine
kinase (SpyM3\_0171), and putative cell wall hydrolase (SpyM3\_0731,
SpyM3\_0922, SpyM3\_1208).  Virulence genes encoding putative nucleases with a
secretion signal peptide were also found to have had experienced horizontal gene
transfer: genome-encoded streptodornase (SDEG\_0541), and prophage-associated
deoxyribonuclease (SDEG\_1103).  Protein F2 like fibronectin-binding protein
(SpyM3\_0104), an adhesin and invasin, was inferred to horizontally
transferred.
